Setting parameters for biological~models with ANIMO

### 2014-04-04 Presentazione/Presentazione_SynCoP.tex
Added more on parameter settings

\documentclass{beamer}
\usetheme{ut}
\usepackage{pgf}
%\usepackage{slashbox}
\usepackage{calc}
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

\frame{\frametitle{All good and nice, but\dots}
{\Huge How to find the parameters?}\\
\begin{itemize}
  \item<1-> Insert parameters manually\only<1-1>{\\\includegraphics[scale=0.3]{images/scenario_parameter_dialog}}
  \item<2-> Compare model versions \emph{subtracting} their activity graphs\only<2-2>{\\\begin{center}\begin{minipage}[c][\graphASize]{\textwidth}\begin{center}\graphA {\Huge -} \includegraphics[scale=0.13]{images/feedback_params_graphD}\end{center}\end{minipage}\xdash[12em]\\ \includegraphics[scale=0.1]{images/difference_network_d-a_larger}\end{center}}
\end{itemize}

@@

\frame{ \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Cytoscape $\rightarrow$ static representation,\\Timed Automata $\rightarrow$ dynamic behaviour
    %\item User interface based on Cytoscape: keep the complexity of Timed Automata \emph{under the hood}
    \item<2-> Discrete levels and simplified scenarios $\rightarrow$ few parameters
    %\item Underlying Timed Automata model provides formal foundations
    \item<3-> Modelling with ANIMO:\\
      \begin{itemize}
          \item outline pathway topology
          \item select reaction parameters
          \item compare model with existing data
          \item repeat until satisfied
      \end{itemize}
    \item<4-> ANIMO allows biologists to draw network ``sketches''
    \item<5-> Complex models:\\
      \begin{itemize}
	\item perform \emph{in silico} experiments
	\item use model checking
      \end{itemize}
  \end{itemize}
}

\frame[t]{ \frametitle{Future work}
  \begin{itemize}
%     \item Advance the development of user interface
%     \item Model checking queries:\begin{itemize} \item user-friendly (templates)
%     \item probabilistic (statistical model checking)
%     \end{itemize}
    \item<1-> Use \emph{in-silico experiments} to infer biological hypotheses which can be verified through in-vitro experiments
      \only<1-1>{\vspace{1ex}\\\includegraphics[width=0.95\textwidth]{images/large_model_sox9_runx2}}
    \item<2-> High-performance analysis of large models: multicore and distributed model checking
    \item<3-> Get a model by feeding data (proposal to NWO):\begin{itemize}
      \item Automata Learning to infer network topology from experimental data
      \item sensitivity analysis and parameter fitting
    \end{itemize}
    \item<4-> Abstraction techniques to deal with large models
  \end{itemize}
}

>>>>

\documentclass{beamer}
\usetheme{ut}
\usepackage{pgf}
%\usepackage{slashbox}
\usepackage{array}
\usepackage{calc}
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

\frame{\frametitle{All good and nice, but\dots}
{\Huge How to find the parameters?}\\
\begin{itemize}
  \item<1-> Insert parameters manually\only<1-1>{\\\includegraphics[scale=0.3]{images/scenario_parameter_dialog}}
  \item<2-> Compare model versions \emph{subtracting} their activity graphs%\only<2-2>{\\\begin{center}\begin{minipage}[c][\graphASize]{.8\textwidth}\begin{center}\graphA \topskip0pt\vspace*{\fill}{\Huge -}\vspace*{\fill} \includegraphics[scale=0.13]{images/feedback_params_graphD}\end{center}\end{minipage}\ \\\xdash[15em]\\ \includegraphics[scale=0.1]{images/difference_network_d-a_larger}\end{center}}
\only<2-2>{\\\begin{center}\begin{tabular}[c]{m{\graphASize}m{1em}m{\graphASize}}\graphA & \begin{center}{\Huge -}\end{center} & \includegraphics[scale=0.13]{images/feedback_params_graphD}\end{tabular}\ \\\xdash[15em]\\ \includegraphics[scale=0.1]{images/difference_network_d-a_larger}\end{center}}
   \item<3-> Perform automatic parameter scans\only<3-3>{\\\includegraphics[width=.8\textwidth]{images/parameters_results_with_bars}}
\end{itemize}

@@

\frame{ \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Cytoscape $\rightarrow$ static representation,\\Timed Automata $\rightarrow$ dynamic behaviour
    %\item User interface based on Cytoscape: keep the complexity of Timed Automata \emph{under the hood}
    \item<2-> Discrete levels and simplified scenarios $\rightarrow$ few parameters
    %\item Underlying Timed Automata model provides formal foundations
    \item<3-> Modelling with ANIMO:\\
      \begin{itemize}
          \item outline pathway topology
          \item select reaction parameters
          \item compare model with existing data
          \item repeat until satisfied
      \end{itemize}
    \item<4-> ANIMO allows biologists to draw network ``sketches''
%    \item<5-> Complex models:\\
%      \begin{itemize}
%	\item perform \emph{in silico} experiments
%	\item use model checking
%      \end{itemize}
     \item<5-> Parameter choice:\\
	\begin{itemize}
		\item manual settings, choice of qualitative parameters
		\item comparison of different model versions
		\item parameter sweeps
	\end{itemize}
  \end{itemize}
}

\frame[t]{ \frametitle{Future work}
  \begin{itemize}
%     \item Advance the development of user interface
%     \item Model checking queries:\begin{itemize} \item user-friendly (templates)
%     \item probabilistic (statistical model checking)
%     \end{itemize}
   \item<1-> Deal with parameter sensitivity and model \emph{robustness}\only<2-3,5->{\\Typically, biological networks are robust: smaller parameter variations do not change the behavior of the network}
    \item<3-> Get a model by feeding data: Automata Learning
    \item<4-> Use \emph{in-silico experiments} to infer biological hypotheses which can be verified through in-vitro experiments
      \only<4-4>{\vspace{1ex}\\\includegraphics[width=0.9\textwidth]{images/large_model_sox9_runx2}}
    %\item<3-> High-performance analysis of large models: multicore and distributed model checking
     % \item sensitivity analysis and parameter fitting
    %\end{itemize}
    \item<5-> Abstraction techniques to deal with large models
  \end{itemize}
}
